---
title: "Day 6"
author: "Julia Romanowska"
---

```{r setup, include = FALSE}
library(tidyverse)
library(here)
library(emoji)
```

## Data

This is the stream of characters that needs to be proceed in sequential
manner to find the _start-of-packet marker_. This is indicated by a sequence
where all four characters are different.

```{r}
(
  stream_input <- scan(
    here("DATA", "2022", "input_day_06.txt"),
    what = "character"
    ) %>%
    strsplit(split = "", fixed = TRUE) %>%
    unlist()
) %>% head(10)
```

# Part one

## Functions

I will need to check at each item:

- whether there are enough characters to create a _marker_ (i.e., min.4) and
- whether all last 4 are different

```{r}
pop_first <- function(cur_read){
  return(cur_read[-1])
}

push_last <- function(cur_read, new_char){
  return(c(cur_read, new_char))
}

update_cur_read <- function(cur_read, new_char){
  new_read <- pop_first(cur_read)
  new_read <- push_last(new_read, new_char)
  return(new_read)
}

check_uniqueness <- function(cur_read){
  n_unique <- length(unique(cur_read))
  return(n_unique == length(cur_read))
}
```

## Process input

```{r}
cur_position <- 4
cur_read <- stream_input[1:cur_position]
```

```{r find_marker}
while (!check_uniqueness(cur_read) | cur_position == length(stream_input)) {
  cur_position <- cur_position + 1
  new_char <- stream_input[cur_position]
  
  cur_read <<- update_cur_read(cur_read, new_char)
}
```

::: {.card .border-success .mb-3}
::: {.card-text}
The number of characters to be processed before the first start-of-packet
marker is detected is `r cur_position`! `r unname(emoji_name["partying_face"])`
:::
:::


# Part two

## Detect _start-of-the-message marker_

The _start-of-the-message marker_ consists of **14 distinct characters**!

```{r}
cur_position <- 14
cur_read <- stream_input[1:cur_position]
```

```{r find_marker}
```

::: {.card .border-success .mb-3}
::: {.card-text}
The number of characters to be processed before the first start-of-message
marker is detected is `r cur_position`! `r unname(emoji_name["partying_face"])`
:::
:::

