---
title: "Day 2"
author: "Julia Romanowska"
---

```{r setup, include = FALSE}
library(tidyverse)
library(here)
library(emoji)
```

## Data

We have a special coding here - let's write a dictionary for that:

- opponents codes:

```
A = "rock", B = "paper", C = "scissors"
```

- my codes:

```
X = "rock", Y = "paper", Z = "scissors"
```

..but the important thing is that each of these give various points, which we
can basically treat as factor-levels:

```{r}
(games_in <- read_delim(
  here("DATA", "input_day_02.txt"),
  delim = " ",
  col_names = FALSE,
  skip_empty_rows = FALSE,
  show_col_types = FALSE
))
games_fct <- games_in %>%
  rename(opponent = X1, me = X2) %>%
  mutate(across(.fns = as.factor))
```

### Explore

```{r}
skimr::skim(games_fct)
levels(games_fct$opponent)
levels(games_fct$me)
```

`r unname(emoji_name["check_mark"])` OK, no missingness and no strange values.

## Check who wins per round

I get `0` if I lose, `3` if it's a draw, and `6` if I win.

```{r}
#' returns points for player1
check_rules <- function(player1, player2){
  player1_lose <- 0
  player1_draw <- 3
  player1_win <- 6

  all_diff <- player1 - player2
  map_dbl(all_diff, function(cur_diff){
    if(cur_diff == 0){ # draw
      return(player1_draw)
    }
    
    if(abs(cur_diff) == 2){ # one of them has 'rock', one 'scissors'
      if(cur_diff < 0){
        return(player1_win)
      }
      
      return(player1_lose)
    }
    
    if(cur_diff > 0){ # check who has higher value
      return(player1_win)
    }
    
    return(player1_lose)
  })
}

(games_fct <- games_fct %>%
  mutate(points = check_rules(as.numeric(me), as.numeric(opponent))))
```

## Get total points

In addition to points gained by drawing or winning a round, I get points for
what I chose (levels of factor).

```{r}
total_score <- games_fct %>%
  mutate(total_per_round = as.numeric(me) + points) %>%
  summarise(grand_total = sum(total_per_round))
```

::: {.card .border-success .mb-3}
::: {.card-text}
The total score is `r sprintf("%d", total_score$grand_total)`!
:::
:::
