---
title: "Day 3"
author: "Julia Romanowska"
---

```{r setup, include = FALSE}
library(tidyverse)
library(here)
library(emoji)
```

## Data

Each line is a list of items in one rucksack.

```{r}
(rucksack_data <- read_table(
  here("DATA", "input_day_03.txt"),
  col_names = FALSE
))
```

## Find the repeating item

First, we need to find the middle of each string and divide.

```{r}
rucksack_divided <- map(
  rucksack_data$X1,
  function(item_list){
    split_at <- str_length(item_list)/2
    out_strings <- c(
      str_sub(item_list, start = 1, end = split_at),
      str_sub(item_list, start = split_at + 1, end = -1)
    )
    return(out_strings)
  }
)
rucksack_divided[[1]]
```

Next, compare the strings in each compartment to find the repeating item for
each rucksack.

```{r}
repeating_items <- map(
  rucksack_divided,
  function(compartments){
    unique_contents <- map(
      compartments,
      function(item_list){
        strsplit(item_list, split = "") %>%
          unlist() %>%
          unique()
      }
    )
    return(Reduce(f = intersect, x = unique_contents))
  }
)
repeating_items[[1]]
```

Checking - do I get only one item per rucksack?

```{r}
lengths(repeating_items) %>%
  unique()
```

`r unname(emoji_name["check_mark"])` Yes, great!

## Priority of items

Lowercase letters have priorities from 1 through 26, uppercase letters - from 27
through 52.

```{r}
priority_low <- str_locate(
  paste0(letters, collapse = ""), letters
)[, "start"]
priority_low <- tibble(
  letter = letters,
  prior = priority_low
)
priority_low

priority_up <- str_locate(
  paste0(LETTERS, collapse = ""), LETTERS
)[, "start"] + 26
priority_up <- tibble(
  letter = LETTERS,
  prior = priority_up
)
priority_up

priority_all <- bind_rows(
  priority_low, priority_up
)
```

Now, find the sum of the priorities of repeated items.

```{r}
total_priority <- tibble(
  repeat_item = unlist(repeating_items)
  ) %>%
  left_join(
    priority_all,
    by = c("repeat_item" = "letter")
  ) %>%
  summarise(sum_priority = sum(prior))
```


::: {.card .border-success .mb-3}
::: {.card-text}
The total priority is `r sprintf("%d", total_priority$sum_priority)`! `r unname(emoji_name["partying_face"])`
:::
:::
